\textbf{Master's Thesis} \\ Correct Projection Mapping on - and Interaction Techniques with - Cloth-based Deformable Surfaces

### 2014-08-03 MSc Report/MScThesis_DanAvram.tex
Added a lot of stuff...running out of time though...

@@

%\author{Dan-Adrian Avram}
%\author{\textbf{Academic Supervisors:} Kasper Hornbaek and Esben Pedersen}

\begin{document}

%\maketitle
\includepdf[pages={-}]{documents/MScFrontPage.pdf}
\thispagestyle{empty}
\newpage
\begin{center}
{\LARGE \textbf{Acknowledgements}}\\
\bigskip
\bigskip
...
\end{center}

@@

In recent years, natural user interfaces have become more and more common. These interfaces are very easy to learn by new users, since they rely on gestures or actions which are already known, unlike physical input devices which have a learning curve and also affect the user's experience, since they have to rely on external, artificial input methods. One of the important developments is the growing research around deformable displays. These displays allow the users to interact with an application as they would do in the real world, by directly manipulating the virtual world with their hands. These types of interfaces build upon the touch paradigm, adding an additional layer of flexibility, allowing gestures to be performed in a 3D space. Unlike mid-air methods, these deformable displays also provide haptic feedback while interacting, which can be useful for certain applications.

@@

While research has been  done in this area, to my knowledge, the problem is still not completely solved. Although the issue was addressed, for example, in \textit{The Deformable Workspace}, they rely on a custom image processor. Further, they only show some images and videos which compare the correction with a normal image, but they have not done a study to determine how users perceive the compensation. In this project, the aim was to build a distortion correction prototype using off the shelf components. Moreover, an empirical study has been done to determine if participants thought that the compensation is working and if they prefer it over a regular method of displaying the image.Originally, the plan was to also create some dynamic applications on which users can test different interaction techniques. However, the distortion problem was more complex than originally thought and, due to several issues, no time was left to develop this part.

Unfortunately, the system has a number of limitations. The main issue is that it has not been calibrated, but manually aligned, which means that the compensation is offset due to misalignments of the physical setup. Several projector-camera calibration techniques were tried, but a workable solution was not employed in the alloted time. Moreover, even though two methods of noise reduction were applied(spatial and temporal averaging), dynamic artifacts were still present in the final image, albeit less pronounced. Finally, another issue is that there is some observable latency from the time of interaction, until the image compensates. Although the image has been updated at around 30 frames per second(\textbf{fps}) with the noise reduction techniques applied, and around 60 fps without.

These issues all constitute candidates for future work. Finally, the compensation only works well from one viewpoint. In order to fix this issue, the user must be tracked and the image has to be rotated depending on the viewpoint.

Although the prototype has several important limitations, the experiment results indicate both that the compensation is working(to some extent) and that the compensated method was almost universally preferred. For this setup, the results convey that the compensation worked best in the center of the image, because other distortions were introduced on the sides, due to the compensation offset. This issue can be fixed by calibrating the system.

@@

\renewcommand{\thepage}{\roman{page}}% Roman numerals for page counter
\setcounter{page}{1}

\section*{Glossary}\mbox{}\\

\begin{abbreviations}
\item[CPU] Central Processing Unit
\item[DMA] Direct Memory Access
\item[FOV] Field of View
\item[fps] Frames Per Second
\item[GPU] Graphics Processing Unit 
\item[MVP] Model View Projection matrix
\item[PTM] Projective Texture Mapping
\item[TOF] Time of Flight 
\item[VBO] Vertex Buffer Object 
\end{abbreviations}
\newpage
\tableofcontents

\newpage
\listoffigures
\newpage
\listoftables 
\newpage
\renewcommand{\thepage}{\arabic{page}}% Arabic numerals for page counter
\setcounter{page}{1}
\section{Introduction}

@@

Moreover, a video that demonstrates the compensation for all scenarios used in the empirical study, as well as other important aspects of the system is available on the included DVD. In addition,the full source code is also available. These resources have also been made available online. Finally, the user study documents and thesis resources can be found in the appendices.

@@

In this chapter, relevant research and related works are highlighted, mainly in relation to deformable user interfaces, tracking and projection techniques on deformable surfaces, interaction techniques with deformable displays, 3D projection and the design of affordances in order to convey information or suggest particular interaction methods. Investigating related research will help in finding solutions to projection issues with deformable surfaces. Moreover, this will also be useful in order to analyze, compare and contrast different approaches to tracking as well as correcting the projection distortion. Further, existing research on interaction techniques and affordances with deformable surfaces will provide a starting point to test, design and implement relevant interaction scenarios for the prototype. Finally, work related to stereoscopic and autostereoscopic projections will be consulted, in order to arrive at a sensible projection technique that can be used in the present case.

\subsection{Tracking and Projection on Deformable Surfaces}

In this section, literature related to projections on deformable surfaces as well as possible distortion correction and tracking techniques techniques will be presented.\\

As a starting point, it makes sense to consider existing approaches of projection onto complex or non-planar physical shapes. As discussed in \cite{daalsgard11}, calibration is needed in order to correct distortions arising from a difference between the digital and the physical object. The study deals with projection on several physical objects, including a complex, organic, non-planar shape, \textbf{Holger the Dane}, as shown in figure \ref{fig:holger}.

@@

Although the information about the necessity of device calibration as well as the projection on a complex surface provide some insights into the projection distortion issue at hand, the study was only concerned with projection on static objects and thus, does not address problems arising from dynamic surface deformation, where the projection would need to be adapted in real-time according to the deformable surface's depth information.\\

\textbf{FlexPad}(figure \ref{fig:flexpad}) is a real-time interactive system with a highly deformable hand-held paper display. A projector is used to display the visuals, while tracking the display surface with a depth camera provides the required depth information for detecting changes in the surface. Further, as described in \cite{steimle13}, any 2D image from Microsoft Windows can be projected and correctly warped unto the display.

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/Flexpad.PNG}
    \caption{Interaction with Flexpad. Courtesy of \cite{steimle13}}
    \label{fig:flexpad}
\end{figure}

@@

After obtaining a deformation model, the tracking problem can be defined as finding parameters of the model such that, when synthesized as a depth image, they match the input depth image best(Analysis by Synthesis or AbS). This represents an optimization problem of finding the best parameter vector. The CMA-ES optimization scheme is used because it is a global optimization algorithm and it applies a smart distribution scheme, minimizing function evaluations needed to find an optimum. More information about these tracking methods will be presented below.

The primary limitations of the FlexPad come from those of the Kinect depth camera not being able to track very sharp bends, however these limitations could be overcome by installing several depth sensors or a more powerful sensor. The hand occlusion issue is not particularly relevant to the current project, since the depth camera will be installed in the back of the display, while interaction shall be from the front  - thus completely removing this problem.

As discussed in \cite{jordt12}, the deformation model can use NURBS surfaces, which can be easily handled via 3D control points, while being able to approximate every surface with arbitrary accuracy. These are widely used in computer graphics and can reduce the high dimensional space of all possible triangle mesh deformations.

The first step is to create an arbitrary NURBS surface using control points and knot vectors, using the algorithm discussed in \cite{jordt12}. Next, once we have a triangle mesh(which can be supplied from a modeling program or computed via OpenGL), we can approximate the mesh with a NURBS surface. The NURBS functions must be fitted to the mesh by minimizing the distance between the surface face function and the vertices of the mesh, or, in other words, finding control points for a given set of vertices, that minimize the sum over all squared distances. Mesh registration can then be performed by associating each 3D vertex coordinates to the previously generated NURBS surface. After this step, the mesh can be translated, rotated and deformed by displacing the control points of the NURBS surface function. Now we can apply an AbS method with a CMA-ES optimization scheme.\\

\textbf{DeforMe} is a projection-based mixed reality (MR) technique, which provides a method for augmenting deformable surfaces with deformation rendering graphics, as described in [4]. Projected graphics are realistically deformed to match the deformed surfaces. Mixed reality creates the illusion of a virtual modification of the physical surfaces. In this study, a flexible surface is used to allow the user to interact with the projected graphics while obtaining tactile feedback. Further, they present some drawbacks of existing tracking methods such as measuring surface deformation via a depth camera, which is not applicable to tangential deformations that can appear on a surface in a horizontal direction during interactions (e.g.: touching, pulling, etc.). To summarize, DeforMe is a projection-based MR system which allows projected graphics to be deformed to the deformation of the real object, while performing real-time estimations of the tangential deformation flows with a small number of feature points. Various deformable materials can be used without prior knowledge of their mechanical properties or deformation models. However, since depth information can be measured for a cloth-based deformable display (with the material properties being known), it is unclear if there is an immediate advantage of this method, for the current project, over the tracking techniques employed in the FlexPad system.\\

@@

They extract depth using imperceptible structured light (projecting binary-coded patterns that are not detected by the visible eye and recording the images with video cameras) - this process can be replaced in the current project by the use of an off-the-shelf depth camera. Further, they describe how, by knowing the object's surface, the viewer's location, projector calibration parameters and the contents to be projected, they can render on irregular 3D surfaces by the use of conventional 3-D methods(e.g.: projective texture mapping), in real-time. Finally, they state that their system can be used to detect display surface changes at non-interactive rates. However, due to a large increase in computing power in recent years, and the possibility to write graphics programs that run directly on the hardware(e.g.: shader programs), similar techniques can be used today, to detect display surface changes in real-time.\\

@@

In [5] they also provide a method to compensate for projection distortion, given that both the calibration parameters and the shape of the deformed screen are known. An observation model in human vision is required to solve this issue, in this case, perspective projection. The technique involves pre-warping the image by shifting each projected point. Further, they claim that projective texture mapping can be used to calculate and render the warped image to be projected in real-time, efficiently and easily. A limitation of the previous approach is that an assumption has been made that the user’s eyes are fixed. However, a viewpoint-based projection should be employed to accommodate realistic scenarios - the virtual space must be rotated and transformed according to the change of viewpoint.\\

As a final note, we can observe that the techniques used to project onto irregular objects and of correcting the visuals are similar in \cite{watanabe08} and \cite{raskar98}: both extract depth information and form a polygonal model of the target object, they use the same rendering technique(projective texture mapping) and require the same calibration parameters.

\newpage
\section{Prototype Design}

Both the projector and the depth camera will be fixed in place relative to the deformable screen. Further, the projected visuals will be interactive applications that need to be rendered continuously. Finally, due to the fact that the projected image will be warped (distorted) according to the screen deformation, an algorithm is needed to shift the projected image pixels corresponding to the display surface’s depth data.

@@

Perhaps another solution would be to use some kind of spatial distortion, similar to the approach used for the \textbf{JellyLens} (see \cite{pindat12}). The focus + context lens(uses two levels of detail to connect a magnified region to the selected context)  can dynamically adapt to the shape of the objects of interest. As we can see in figure  \ref{fig:JellyLens}, the object of interest adapts its shape(is magnified), while the context is preserved - surrounding objects are almost untouched by the applied distortion.

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/JellyLens.PNG}
    \caption{a) Only a portion of the object can be magnified with a small fisheye lens; b) A large fisheye magnifies almost the entire object, but at the cost of added distortion for sorrounding objects; c) JellyLens - adapting the relevant information of the object, while preserving the surrounding areas. Courtesy of \cite{pindat12}}
    \label{fig:JellyLens}
\end{figure}

For this project, a solution similar to that employed in \cite{watanabe08} has been implemented. In order to get a better overview and understanding of the whole process required to correct the distortions, I have created a system pipeline(see figure \ref{fig:Pipeline}) that describes the whole process, at a high-level. Each part will be described individually in subsequent chapters.

@@

Further, the application contents must be displayed in such a way that they distort with the object. Like with most problems in computer graphics, there is more than one solution. One could, for example, use raytracing to directly project rays of light on the object(also mentioned by the authors of \cite{raskar98}. However, this approach is known to be slow and may not run in real-time. Alternatively, there exists a technique which has been tried and tested before, known in computer graphics, which projects a texture onto an object as from a slide projector. This technique can also be implemented in real-time on modern hardware, and, since the texture is projected from a virtual projector in the 3D scene, it will update with the distortions corresponding to the deformation on the object.\\

Finally, we need to make sure that our virtual camera corresponds, as closely as possible to the area we project onto with our physical display. Further, we also need to place our virtual projector, which also requires information about where the mesh(3D model) is placed, and how it is oriented. In addition to position and orientation we must also know the field of view(FOV), or the \textit{"zoom level"} of our camera(as the other parameters depend on the FOV). In addition to the calibration required above, one also needs to manually(physically) align the projector so that if correctly fills the deformable display area. Furthermore, in case of manual calibration, the depth camera must also be placed as orthogonal as possible to the deformable surface area, so that it is easy to obtain acceptable parameters for the virtual camera and projector described above.

At this point, we basically have a solution to the distortion correction problem. The only remaining issue is to update the application contents dynamically, by capturing window/screen contents. This can be accomplished by using textures.

In the next sections, each of the above will be described in greater detail.

@@

Secondly, we need to compute the modelview matrix(see figure \ref{fig:MVMatrix}, for the structure of such a matrix). This matrix has information that relates to the position of the model(3D object space) as well as the camera(eye space). In this case, we essentially need to compute the view matrix from the extrinsic parameters, so that our virtual camera correctly points to the deformable display area on which our target contents are projected. \textbf{Extrinsic parameters} represent the \textbf{location} and the \textbf{orientation} of the projector/camera with respect to a world reference frame. For example, if we choose the world origin to be the position of the depth camera, then we can compute the rotation and translation of the projector with respect to the depth camera by finding the rotation matrix R and translation coefficients T. To obtain camera coordinates from world coordinates we simply need to multiply a position by R and add the translation T.

@@

\subsubsection{Approaches to Calibration}

@@

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/GeneralPinholeModel.PNG}
    \caption{General Pinhole Camera Model}
    \label{fig:PinholeCameraModel}
\end{figure}

@@

\subsubsection{Other Methods for 2.5D Triangulation}

The naive triangulation, by connecting the triangles without considering some measure of how natural it looks, may not provide the intuitively \textit{best-looking} results. If the triangulation is not acceptable, it can be improved with a technique such as \textbf{Delaunay Triangulation}(explained in detail in \cite[Chapter~9]{berg08}), where triangulations are ranked by comparing their smallest angle(small angles, result in \textit{skinniness}). The issue is shown in figure \ref{fig:MountainRidgeNarrowValley} where two triangulations of the same point set are shown. From the heights of the sample points, it seems as though the points were taken from a \textit{mountain ridge}. If an edge is flipped to obtain very small angles for our triangles, we get what looks as a \textit{narrow valley}(as you can see, the height value is now very small). The latter is a suboptimal triangulation and can be improved. Delaunay triangulation aims to find an optimal triangulation, given the fact that, for any set of points, there is a finite amount of possible triangulations.

\subsubsection{Marching Cubes on GPU}

One of the classic computer algorithms for surface reconstruction is the \textit{Marching Cubes}  algorithm. An attempt was made to implement this algorithm. In order to run it on a GPU, one can use a \textit{geometry shader} or\textit{ histogram pyramids} with a regular vertex and fragment shader. The attempt consisted of porting an existing implementation of the geometry shader for fixed-pipeline OpenGL to modern OpenGL. Unfortunately, I did not get any geometry as output after running the algorithm, but there were also no errors. The problem is that I could not debug the shaders, since the tools I have tried did not work - some because of no support for modern OpenGL(3.0+) like \textit{glslDevil}, and others did not work on my laptop because it uses the \textit{nVidia Optimus} technology, as is the case for the \textit{nVidia nSight} debugging tool for \textit{Microsoft Visual Studio}. As this was taking too much time, I decided to go for a simpler approach, that should still prove appropriate for the real-time 3D reconstruction of the cloth, described in section \ref{sec:TriangleStripTriangulation}. Nevertheless I will briefly describe the algorithm in the following section, as well as how it can be implemented on a GPU. This may be useful to look into, as an extension to the current project, if a higher quality mesh is desired(this method works well for 3D objects in general, and is not restricted to 2.5D, as is the case for the currently used method).

@@

\paragraph{GPU Implementation}\mbox{}\\

Another approach is to use histogram pyramids, as described in \cite{dyken2007}. The authors state that their algorithm outperforms the known geometry shader approaches and it does not take much more effort to implement. The hierarchical data structured called the \textit{HistoPyramid}, used for data expansion and compaction of 2D textures, is the focus of the previously mentioned paper. The term texel was used to denote single data elements, and the 3D array of voxels has been mapped to a 2D domain.

\paragraph{Implicit Surface Generation}\mbox{}\\

As stated before, in order to use the marching cubes algorithm, one also needs to obtain a scalar value at each point in the data set, and, since we only have a point cloud, a function is needed to provide this - such a dataset is called an implicit surface. These functions were not a focus for this project, as the marching cubes algorithms was not successfully run on datasets that already represented implicit surfaces. However, in \cite{navpreet2013} the Hermite Radial Basis Function(HRBF) is mentioned as a workable method, please consult this resource for more information on the topic. Another function that can determine the scalar field is the Newtonian gravity field equation, please consult \cite{max2013} for further information.

\subsubsection{Greedy Triangulation}

>>>>

\usepackage{fancyhdr}

@@

%\author{Dan-Adrian Avram}
%\author{\textbf{Academic Supervisors:} Kasper Hornbaek and Esben Pedersen}
\pagestyle{fancy}
%\renewcommand{\sectionmark}[1]{\leftmark{#1}}
\fancyhf{}
%\fancyhead[R]{MyName} % predefined ()
%\fancyhead[R]{\rightmark} % 1. sectionname
%\fancyhead[L]{\leftmark} % 1. sectionname

%\fancypagestyle{plain}{%
%  \fancyhf{}%
%  \renewcommand{\headrulewidth}{0pt}%
%}
% Now set the headers
%\fancyhead[LE,RO]{\footnotesize\rightmark}
%\fancyhead[LO,RE]{\footnotesize \leftmark}
\fancyhead[LE,RO]{\footnotesize \ifthenelse{\isodd{\value{page}}}{\rightmark}{\leftmark}}
\fancyfoot[C]{\thepage}
%{\ifthenelse{\isodd{\value{page}}}{\leftmark}{\rightmark}

\begin{document}
\setcounter{tocdepth}{5}
\includepdf[pages={-}]{documents/MScFrontPage.pdf}

@@

In recent years, natural user interfaces have become more and more common. These interfaces are very easy to learn by new users, since they rely on gestures or actions which are already known, unlike physical input devices which have a learning curve. In addition, these artificial input methods also affect the users' experience, since there is an additional layer of abstraction between them and the application. One of the recent important developments, is the growing research around deformable displays. These displays allow the users to interact with an application as they would do in the real world, by directly manipulating the virtual world with their hands. These types of interfaces build upon the touch paradigm, adding an additional layer of flexibility, allowing gestures to be performed in a 3D space. Unlike mid-air methods, these deformable displays also provide haptic feedback while interacting, which can be useful for certain applications.

@@

While research has been  done in this area, to my knowledge, the problem is still not completely solved. Although the issue was addressed, for example, in \textit{The Deformable Workspace}, they rely on a custom, dedicated hardware image processor. Further, they only compare some compensated images and videos with normal ones, but they have not done a study to determine how users perceive the compensation. In this project, the aim was to build a distortion correction prototype using off the shelf components. Moreover, an empirical study has been done to determine if participants thought that the compensation is working and if they prefer it over a regular method of displaying the image.

%Unfortunately, the system has a number of limitations. The main issue is that it has not been calibrated, but manually aligned, which means that the compensation is %offset due to misalignments of the physical setup. Several projector-camera calibration techniques were tried, but a workable solution was not employed in the %alloted time. Moreover, even though two methods of noise reduction were applied(spatial and temporal averaging), dynamic artifacts were still present in the final %image, albeit less pronounced. Finally, another issue is that there is some observable latency from the time of interaction, until the image compensates. Although %the image has been updated at around 30 frames per second(\textbf{fps}) with the noise reduction techniques applied, and around 60 fps without.

%These issues all constitute candidates for future work. Finally, the compensation only works well from one viewpoint. In order to fix this issue, the user must be %tracked and the image has to be rotated depending on the viewpoint.

Although the prototype has some limitations, the experiment results indicate both that the compensation is working(to some extent) and that the compensated method was almost universally preferred. For this setup, the results convey that the compensation worked best in the center of the image, because other distortions were introduced on the sides, due to a compensation offset. This issue can be fixed by calibrating the system.

@@

\renewcommand{\thepage}{\roman{page}}% Roman numerals for page counter
\setcounter{page}{1}
\thispagestyle{empty}
\section*{Glossary}\mbox{}\\

\begin{abbreviations}
\item[CPU] Central Processing Unit
\item[DMA] Direct Memory Access
\item[DOF] Degrees of Freedom
\item[FOV] Field of View
\item[fps] Frames Per Second
\item[GPU] Graphics Processing Unit 
\item[MVP] Model View Projection matrix
\item[NURBS] Non-Uniform Rational B-Spline
\item[QUIS] Questionnaire For User Interaction Satisfaction
\item[PTM] Projective Texture Mapping
\item[SDK] Software Development Kit
\item[TOF] Time of Flight 
\item[VAO] Vertex Array Object 
\item[VBO] Vertex Buffer Object 
\end{abbreviations}
\newpage
\tableofcontents

\newpage
\listoffigures
\newpage
\listoftables

\newpage
\thispagestyle{empty}
\renewcommand{\thepage}{\arabic{page}}% Arabic numerals for page counter
\setcounter{page}{1}
\section*{Preface}
%\begin{center}
%{\LARGE \textbf{Preface}}\\
%\end{center}
%\bigskip
%\bigskip
This report is the written result of my Master's thesis project at the Department of Computer Science at the University of Copenhagen.

I would like to thank my academic advisor, Kasper Hornb\ae k as well as my co-advisor, Esben Pedersen, for guidance and feedback throughout the entire project. In addition, I would also like to thank Kenny Erleben for help on computer graphics issues.

Special thanks go to all who participated in the pilot and actual studies as well as to several members of the Human-Centered Computing group, who provided valuable feedback. Finally, I would like to thank my family and friends for their continued support.

\newpage
\section{Introduction}

@@

Moreover, a video that demonstrates the compensation for all scenarios used in the empirical study, as well as other important aspects of the system is available on the included DVD or on-line(see appendix \ref{appendix:A}). In addition,the full source code is also available. These resources have also been made available online. Finally, the user study documents and thesis resources can be found in the appendices.

@@

In this chapter, relevant research and related works are highlighted, mainly in relation to deformable user interfaces, tracking and projection techniques on deformable surfaces. Investigating related research will help in finding solutions to projection issues with deformable surfaces. Moreover, this will also be useful in order to analyze, compare and contrast different approaches to tracking as well as correcting the projection distortion.

\subsection{Projection and Tracking on Deformable Surfaces}

\subsubsection{Projection on Non-Planar Surfaces}
As a starting point, it makes sense to consider existing approaches of projection onto complex or non-planar physical shapes. As discussed in \cite{daalsgard11}, calibration is needed in order to correct distortions arising from a difference between the digital and the physical object. The study deals with projection on several physical objects, including a complex, organic, non-planar shape, \textbf{Holger the Dane}, as shown in figure \ref{fig:holger}.

@@

Although the information about the necessity of device calibration, as well as the projection on a complex surface, provide some insights into the projection distortion issue at hand, the study was only concerned with projection on static objects. Thus, it does not address problems arising from dynamic surface deformation, where the projection would need to be adapted in real-time, according to the deformable surface's depth information.\\

Next, \textbf{DeforMe} is a projection-based mixed reality (MR) technique, which provides a method for augmenting deformable surfaces with deformation rendering graphics, as described in [4]. Projected graphics are realistically deformed to match the deformed surfaces. Mixed reality creates the illusion of a virtual modification of the physical surfaces. In this study, a flexible surface is used to allow the user to interact with the projected graphics while obtaining tactile feedback. To summarize, DeforMe is a projection-based MR system which allows projected graphics to be deformed to the deformation of the real object. Various deformable materials can be used without prior knowledge of their mechanical properties or deformation models. However, the aim here is to make the graphics adapt realistically to the surface deformation, while the objective in our current project is to correct all deformations, so that the image looks like on a flat display.

\subsubsection{Tracking Techniques on Deformable Objects}

 In the \textbf{DeforMe} project, the authors present some drawbacks of existing tracking methods such as measuring surface deformation via a depth camera, which is not applicable to tangential deformations that can appear on a surface in a horizontal direction during interactions (e.g.: touching, pulling, etc.). However, we are not interested to model the deformations of the shape realistically, but to correct them, meaning that a depth sensor is adequate to capture the deformation.\\

%However, since depth information can be measured for a cloth-based deformable display (with the material properties being known), it is unclear if there is an %immediate advantage of this method, for the current project, over the tracking techniques employed in the FlexPad system.\\

\textbf{FlexPad}(see \cite{steimle13}), is a real-time interactive system with a highly deformable hand-held paper display. A projector is used to display the visuals, while tracking the display surface with a depth camera provides the required depth information for detecting changes in the surface. Further, as described in the paper, any 2D image from Microsoft Windows can be projected and correctly warped unto the display.

%\begin{figure}[hbtp]
%    \centering
%    \includegraphics[width=0.8\textwidth]{figures/Flexpad.PNG}
%    \caption{Interaction with Flexpad. Courtesy of \cite{steimle13}}
%    \label{fig:flexpad}
%\end{figure}

@@

After obtaining a deformation model, the tracking problem can be defined as finding parameters of the model such that, when synthesized as a depth image, they match the input depth image best(Analysis by Synthesis or AbS). This represents an optimization problem of finding the best parameter vector. More information about this tracking method is available in \cite{steimle13}.

The primary limitations of the FlexPad come from those of the Kinect depth camera not being able to track very sharp bends, however these limitations could be overcome by installing several depth sensors or a more powerful sensor. The hand occlusion issue is not particularly relevant to the current project, since the depth camera was installed in the back of the display, while interaction was done from the front - thus completely removing this problem.

As discussed in \cite{jordt12}, the deformation model can use NURBS surfaces, which can be easily handled via 3D control points, while being able to approximate every surface with arbitrary accuracy. These are widely used in computer graphics and can reduce the high dimensional space of all possible triangle mesh deformations(only the 25x25 vertex plane is used).

The first step is to create an arbitrary NURBS surface using control points and knot vectors, using the algorithm discussed in \cite{jordt12}. Next, once we have an initial triangle mesh(which can be supplied from a modeling program or computed via OpenGL, by reconstructing the object in a static video frame), we can approximate the mesh with a NURBS surface. The NURBS functions must be fitted to the mesh by minimizing the distance between the surface face function and the vertices of the mesh, or, in other words, finding control points for a given set of vertices, that minimize the sum over all squared distances. Mesh registration can then be performed by associating each 3D vertex coordinate to the previously generated NURBS surface. After this step, the mesh can be translated, rotated and deformed by displacing the control points of the NURBS surface function.\\

An alternative to this tracking method is employed in \cite{watanabe08}. Moreover, deformation correction is also implemented, in the previously mentioned paper. Therefore, it is described in the following section.

\subsection{Distortion Correction on Deformable Surfaces}

@@

They extract depth using imperceptible structured light (projecting binary-coded patterns that are not detected by the visible eye and recording the images with video cameras) - this process can be replaced in the current project by the use of an off-the-shelf depth camera. Further, they describe how, by knowing the object's surface, the viewer's location, projector calibration parameters and the contents to be projected, they can render on irregular 3D surfaces by the use of conventional 3-D methods(e.g.: \textbf{PTM}(projective texture mapping)), in real-time. Finally, they state that their system can be used to detect display surface changes at non-interactive rates. However, due to a large increase in computing power in recent years, and the possibility to write custom graphics programs that run directly on the hardware(e.g.: shader programs), similar techniques can be used today, to detect display surface changes in real-time.\\

@@

In [5] they also provide a method to compensate for projection distortion, given that both the calibration parameters and the shape of the deformed screen are known. The technique involves pre-warping the image by shifting each projected point. Further, they claim that PTM can be used to calculate and render the warped image to be projected in real-time, efficiently and easily. A limitation of the previous approach is that an assumption has been made that the user’s eyes are fixed. However, a viewpoint-based projection should be employed to accommodate realistic scenarios - the virtual space must be rotated and transformed according to the change of viewpoint.\\

Perhaps another solution, would be to use some kind of spatial distortion, similar to the approach used for the \textbf{JellyLens} (see \cite{pindat12}). The focus + context lens(uses two levels of detail to connect a magnified region to the selected context)  can dynamically adapt to the shape of the objects of interest. As we can see in figure  \ref{fig:JellyLens}, the object of interest adapts its shape(is magnified), while the context is preserved - surrounding objects are almost untouched by the applied distortion.

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/JellyLens.PNG}
    \caption{a) Only a portion of the object can be magnified with a small fisheye lens; b) A large fisheye magnifies almost the entire object, but at the cost of added distortion for surrounding objects; c) JellyLens - adapting the relevant information of the object, while preserving the surrounding areas. Courtesy of \cite{pindat12}}
    \label{fig:JellyLens}
\end{figure}

\subsection{Chapter Summary}

This review has highlighted some of the main methods of projecting, tracking and correcting distortions for non-planar deformable objects. Further , we could observe that the techniques used to project onto irregular objects and to correct the visuals, are similar in \cite{watanabe08} and \cite{raskar98}: both extract depth information and form a polygonal model of the target object, they use the same rendering technique(PTM) and require the same calibration parameters. Moreover, the NURBS deformation model used in \cite{steimle13}, could be employed as an alternative to the methods described in the two previously mentioned sources.

The existing work on this topic helped to identify the main steps that are required in order to compensate an image and correct it's distortion. Information about alternatives to some of these steps has also surfaced. This literature review has guided my choices through the remaining chapters.

\newpage
\section{Prototype Design and Theoretical Considerations}

Both the projector and the depth camera were fixed in place relative to the deformable screen. Further, the projected visuals would normally be interactive applications for such a display - they need to be rendered continuously. Finally, due to the fact that the projected image was warped (distorted) according to the screen deformation, an algorithm was needed to shift the projected image pixels corresponding to the display surface’s depth data.

@@

For this project, a solution similar to that employed in \cite{watanabe08} has been implemented. This is because it was the only paper where distortion correction in a 3D space was more thoroughly addressed, thus reducing risk. Further, similar techniques could be easily used with our hardware setup. It was unclear whether the other methods, such as the NURBS model used in \cite{steimle13} would be accurate enough for compensation - the deformation model is approximated; if the offset is too large, deformations may not look right.

In order to get a better overview and understanding of the whole process required to correct the distortions, I have created a system pipeline(see figure \ref{fig:Pipeline}) that describes the whole process, at a high-level. Each part will be described individually in subsequent chapters.

@@

Further, the application contents must be displayed in such a way that they distort with the object. Like with most problems in computer graphics, there is more than one solution. One could, for example, use raytracing to directly project rays of light on the object(also mentioned by the authors of \cite{raskar98}). However, this approach is known to be slow and may not run in real-time. Alternatively, there exists a technique which is known in computer graphics, which projects a texture onto an object as from a slide projector(PTM). This technique can also be implemented in real-time on modern hardware. Further, since the texture is projected from a virtual projector in the 3D scene, it will update with the distortions corresponding to the deformation on the object.\\

Finally, we need to make sure that our virtual camera corresponds, as closely as possible to the area we project onto with our physical display. Further, we also need to place our virtual projector, which requires information about where the mesh(3D model) is placed, and how it is oriented. In addition to position and orientation we must also know the field of view(FOV), or the \textit{"zoom level"} of our camera(as the other parameters depend on the FOV). In addition to the calibration required above, one also needs to manually(physically) align the projector so that if correctly fills the deformable display area. Furthermore, in case of manual calibration, the depth camera must also be placed as orthogonal as possible to the deformable surface area, so that it is easy to obtain acceptable parameters for the virtual camera and projector described above. At this point, we basically have a solution to the distortion correction problem. The only remaining issue is to update the application contents dynamically, by capturing window/screen contents. This can be accomplished by using textures. In the next sections, each of the above will be described in greater detail.

@@

Secondly, we need to compute the modelview matrix(see figure \ref{fig:MVMatrix}, for the structure of such a matrix). This matrix has information that relates to the position of the 3D object(model) as well as the camera(eye space). In this case, we essentially need to compute the view matrix from the extrinsic parameters, so that our virtual camera correctly points to the deformable display area on which our target contents are projected. \textbf{Extrinsic parameters} represent the \textbf{location} and the \textbf{orientation} of the projector/camera with respect to a world reference frame. For example, if we choose the world origin to be the position of the depth camera, then we can compute the rotation and translation of the projector with respect to the depth camera by finding the rotation matrix R and translation coefficients T. To obtain camera coordinates from world coordinates we simply need to multiply a position by R and add the translation T.

@@

\begin{figure}[hbtp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/GeneralPinholeModel.PNG}
    \caption{General Pinhole Camera Model}
    \label{fig:PinholeCameraModel}
\end{figure}
\newpage
\subsubsection{Approaches to Calibration}

@@

@@

\subsubsection{Other Methods for Triangulation}

Other methods of triangulating the mesh into either 2.5D or 3D objects will be shortly described in the next sections. An implementation of most of these methods has been attempted. However, the technique described in section \ref{sec:TriangleStripTriangulation} is the one which is currently used. Some of these alternatives were running either too slow, or were taking too long to implement.

\paragraph{Delaunay Triangulation}\mbox{}\\

The naive triangulation, by connecting the triangles without considering some measure of how natural it looks, may not provide the intuitively \textit{best-looking} results. If the triangulation is not acceptable, it can be improved with a technique such as \textbf{Delaunay Triangulation}(explained in detail in \cite[Chapter~9]{berg08}), where triangulations are ranked by comparing their smallest angle(small angles, result in \textit{skinniness}). This tecnhique can be used for 2.5D reconstruction and can obtain an optimal triangulation. For more information please consult appendix \ref{appendix:B-delaunay}.

\paragraph{Marching Cubes}\mbox{}\\

One of the classic computer algorithms for surface reconstruction is the \textit{Marching Cubes}  algorithm. An attempt was made to implement this algorithm. In order to run it on a GPU, one can use a \textit{geometry shader} or\textit{ histogram pyramids} with a regular vertex and fragment shader. The attempt consisted of porting an existing implementation of the geometry shader for fixed-pipeline OpenGL to modern OpenGL. Unfortunately, I did not get any geometry as output after running the algorithm, but there were also no errors. The problem is that I could not debug the shaders, since the tools I have tried did not work - some because of no support for modern OpenGL(3.0+) like \textit{glslDevil}, and others did not work on my laptop because it uses the \textit{nVidia Optimus} technology, as is the case for the \textit{nVidia nSight} debugging tool for \textit{Microsoft Visual Studio}. As this was taking too much time, I decided to go for a simpler approach, that should still prove appropriate for the real-time 3D reconstruction of the cloth, described in section \ref{sec:TriangleStripTriangulation}. Nevertheless I will briefly describe the algorithm in the following section. More information on how it can be implemented on a GPU is available in appendix \ref{appendix:B-mc}. This may be useful to look into, as an extension to the current project, if a higher quality mesh is desired(this method works well for 3D objects in general, and is not restricted to 2.5D, as is the case for the currently used method).

@@

\paragraph{Greedy Triangulation}\mbox{}\\

Another triangulation method that was briefly considered was the greedy projection algorithm described in \cite{Marton09ICRA}. As described in the previously mentioned source as well as in \cite{navpreet2013}, the algorithm grows the mesh from a list of points, until all points have been considered or there are no more valid triangles to be formed - in case of the latter, the algorithm is restarted somewhere in the unconnected area with vertices from a new \textit{seed} triangle.
